---
title: "bird_joins"
format: html
---
## Load Libraries
```{r}
library(tidyverse)
library(here)
library(stringr)
library(dplyr)
```
## Read in data csv
```{r}
bird_observations <- read_csv(here("data","52_pp52_birds_1.csv"))
sites <- read_csv(here("data","52_pp52_sites_1.csv"))
surveys <- read_csv(here("data","52_pp52_surveys_1.csv")) 
taxalist <- read_csv(here("data","52_pp52_taxalist_1.csv"))
```

# Create a subset of bird_observations called birds_subset that only contains observations for birds with species id “BHCO” and “RWBL”, and from sites with site ID “LI-W” and “NU-C”
```{r}
bird_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))
  
```

## Left join practice
Use left join(s) to update birds_subset so that it also *includes sites and taxalist* information. *For each join*, include an *explicit argument saying which variable you are joining by (even if it will just assume the correct one for you)*. Store the updated data frame as birds_left. Make sure to look at the output - is what it contains consistent with what you expected it to contain?
```{r}
birds_left <- bird_subset %>% # using bird_subset data
  left_join(sites, "site_id") %>% # left joining sites data, by matching "site_id" containing the same column in bird_subset
  left_join(taxalist, "species_id") #left joining sites data, by matching "species_id" containing the same column in bird_subset
```

#Full join practice
First, answer: what do you expect a full_join() between birds_subset and sites to contain?
- I would expect for full join- to add all the data from 'sites' dataframe into bird_subset 
# Write code to full_join the birds_subset and sites data into a new object called birds_full. Explicitly include the variable you’re joining by. Look at the output. Is it what you expected?
```{r}
birds_full <- full_join(bird_subset, sites)
```

## Task 2: Data wrangling and visualization with birds
Starting with your birds object, rename the notes column to bird_obs_notes (so this doesn’t conflict with notes in the surveys dataset



