---
title: "string_practice"
format: html
---
```{r}
library(tidyverse)
library(here)
library(stringr)
library(dplyr)
library(lubridate)
```

# Copy all contents of the html table below to your clipboard:
- 
date	building	alarm_message
2020-03-14	Engineering-North	10:02am -- HVAC system down, facilities management alerted
2020-03-15	Bren Hall	8:24am -- Elevator North out of service
2020-04-10	Engineering-South	12:41am -- Fire alarm, UCSB fire responded and cleared
2020-04-18	Engr-North	9:58pm -- Campus point emergency siren, UCPD responded
# Back in your string_practice.Rmd, create a new code chunk
# With your cursor in your code chunk, go up to Addins in the top bar of RStudio. From the drop-down menu, choose ‘Paste as data frame’. Make sure to add code to store the data frame as alarm_report
```{r}
alarm_report<- data.frame(
  stringsAsFactors = FALSE,
              date = c("2020-03-14", "2020-03-15", "2020-04-10", "2020-04-18"),
          building = c("Engineering-North",
                       "Bren Hall","Engineering-South","Engr-North"),
     alarm_message = c("10:02am -- HVAC system down, facilities management alerted",
                       "8:24am -- Elevator North out of service",
                       "12:41am -- Fire alarm, UCSB fire responded and cleared",
                       "9:58pm -- Campus point emergency siren, UCPD responded")
)
```

#Practice working with strings by writing code to update alarm_report as follows (these can be separate, or all as part of a piped sequence):
- Replace the “Engr” with “Engineering” in the building column
- Separate the building column into two separate columns, building and wing, separated at the dash
- Only keep observations with the word “responded” in the alarm_message column
- Separate the message time from the rest of the message by separating at --
- Convert the date column to a Date class using lubridate
```{r}
alarm_report %>% 
  mutate(building = str_replace(building, pattern = "Engineering", replacement = "Engr")) %>% # Replacing the name from Engineering to Engr
  separate(col= building,
           into = c("building", "wing"),
           sep = "-") %>% # separating into building and wing column
  filter(str_detect(alarm_message, pattern = "responded")) %>% 
  separate(col= alarm_message,
           into = "time",
           sep = "am") %>% 
  mutate(date = lubridate::ymd(date))   

```

